---
title: "HW 2 Student"
author: "Andy Ackerman"
date: "10/17/2023"
output: 
  html_document:
    number_sections: true
---

This homework is meant to illustrate the methods of classification algorithms as well as their potential pitfalls.  In class, we demonstrated K-Nearest-Neighbors using the `iris` dataset.  Today I will give you a different subset of this same data, and you will train a KNN classifier.  

```{r, echo = FALSE}
set.seed(123)
library(class)
library(tidyverse)

df <- data(iris) 

normal <-function(x) {
  (x -min(x))/(max(x)-min(x))   
}

iris_norm <- as.data.frame(lapply(iris[,c(1,2,3,4)], normal))

subset <- c(1:45, 58, 60:70, 82, 94, 110:150)
iris_train <- iris_norm[subset,] 
iris_test <- iris_norm[-subset,] 

iris_target_category <- iris[subset,5]
iris_test_category <- iris[-subset,5]


```

#
Above, I have given you a training-testing partition.  Train the KNN with $K = 5$ on the training data and use this to classify the 50 test observations.  Once you have classified the test observations, create a contingency table -- like we did in class -- to evaluate which observations your algorithm is misclassifying.   

```{r}
set.seed(123)

tab <- knn(iris_train, iris_test, cl = iris_target_category, k = 5) %>% # KNN Test using iris_train as the training data, iris_test as the testing data, iris_target_category as the classifer, and classifying based on the class of the 5 nearest neighbors
  table(iris_test_category) # Contingency table

tab

accuracy <- function(x){
  sum(diag(x)/(sum(rowSums(x)))) * 100 # Calculates the accuracy of the model
}
accuracy(tab)
```

#

Discuss your results.  If you have done this correctly, you should have a classification error rate that is roughly 20% higher than what we observed in class.  Why is this the case? In particular run a summary of the `iris_test_category` as well as `iris_target_category` and discuss how this plays a role in your answer.  

```{r}
summary(iris_test_category)
summary(iris_target_category)
```

*The model incorrectly classified 11 virginica plants as versicolor, resulting in a 78% accuracy. This is lower than the in-class version because the training set is not representative of the testing set --- versicolor plants are overrepresented. Therefore, because there are more versicolor plants clustered, higher values of k will cause a point to be more likely to be classified as versicolor.* 

#

Build a github repository to store your homework assignments.  Share the link in this file.  

*STUDENT INPUT*

